---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
 
Loading packages to be used:
```{r}
library(lubridate)
library(dplyr)
library(lattice)
```

As I'm from Brazil, week days labels would change if I didn't set the system locale:
```{r}
Sys.setlocale("LC_TIME","C") 
```


## Loading and preprocessing the data
Loading data:
```{r}
dados<-read.csv("activity.csv", h = T)
```

Transforming date to `POSIXct` class:
```{r}
dados<-transform(dados, date = ymd(date))
```

Wrapping the data frame in a dplyr tbl_df:
```{r}
dados<-tbl_df(dados)
```


## What is mean total number of steps taken per day?
Grouping data by date (using dplyr package):
```{r}
tsdados<-summarise(group_by(dados,date), sum(steps, na.rm = T))
names(tsdados)<-c("date", "sumSteps")
```

Mean and median of total number of steps per day:
```{r}
mean(tsdados$sumSteps)
median(tsdados$sumSteps)
```

So, the mean of total numer of steps per day is `r mean(tsdados$sumSteps)` and it's median is `r median(tsdados$sumSteps)`.

To illustrate all this, here comes a histogram:
```{r}
hist(tsdados$sumSteps, breaks = 30, xlab = "Total number of steps per day", main = NULL)
```


## What is the average daily activity pattern?

Grouping data by interval:
```{r}
daDados<-summarise(group_by(dados,interval), mean(steps, na.rm = T))
names(daDados)<-c("interval","meanSteps")
```

Time series plot of 5-minute intervals and average number of steps per day:
```{r}
plot(daDados$interval, daDados$meanSteps, type = "l", xlab = "5-minute intervals", ylab = "Average number of steps")
```


Interval that contains, on average, the maximum number of steps:

```{r}
daDados[daDados$meanSteps == max(daDados$meanSteps),]$interval
```



So, the interval then interval that contains, on average, the maximum number of steps is ```r daDados[daDados$meanSteps == max(daDados$meanSteps),]$interval```, with an average number of steps = ```r daDados[daDados$meanSteps == max(daDados$meanSteps),]$meanSteps```.

## Imputing missing values
Calculating total number of missing values in the data set:

```{r}
sum(is.na(dados$steps))
```

The total number of missing values (```r "NA" ```) in this data set is ```r sum(is.na(dados$steps))```.


Function for imputing missing values in the data set.
```{r}
imputation<-function (dataframe){
    vetor = vector()
    
    for (linha in 1:nrow(dataframe)){
        passos=dataframe[linha,]$steps
            
        if (is.na(passos)){
            intervalo = dataframe[linha,]$interval
            x = daDados[daDados$interval == intervalo,]$meanSteps
            vetor<-append(vetor, x)
        }
        
        else {
            vetor<-append(vetor, passos)
        }
    }
    
    novoDataframe<-mutate(dataframe, steps = vetor)
    novoDataframe    
}

```


New data set, equal to original, with missing data filled in:
```{r}
dadosComp<-imputation(dados)
```



Grouping data by date (using dplyr package):
```{r}
tsDadosComp<-summarise(group_by(dadosComp,date), sum(steps))
names(tsDadosComp)<-c("date", "sumSteps")
```


Mean and median of total number of steps per day:
```{r}
mean(tsDadosComp$sumSteps)
median(tsDadosComp$sumSteps)
```

So, the mean of total numer of steps per day is `r mean(tsDadosComp$sumSteps)` and it's median is `r median(tsDadosComp$sumSteps)`.

To illustrate all this, here comes a histogram:
```{r}
hist(tsDadosComp$sumSteps, breaks = 30, xlab = "Total number of steps per day", main = NULL)
```

######Falta resposta




## Are there differences in activity patterns between weekdays and weekends?


Function to detect weekend or weekday
```{r}
weekend<- function(data){
    dia<-weekdays(data)
    if (dia == "Saturday" | dia == "Sunday"){
        dia = "weekend"
    }
    else {
        dia = "weekday"
    }
    dia
}
    
```

Creating a new factor variable in the data frame:
```{r}
dadosWeek<-mutate(dadosComp, week = factor(do.call(rbind, lapply(dadosComp$date, weekend))))
```

Grouping data by time interval and weekday factor ("week"), and extracting mean:
```{r}
tsDadosWeek<-summarise(group_by(dadosWeek, week, interval), mean(steps))
names(tsDadosWeek)<-c("week", "interval", "steps")
```

Plotting:
```{r}
xyplot(steps ~ interval | week, data = tsDadosWeek, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")
```


